---
title: "Walk Or Run"
author: "NM"
date: "19 juillet 2017"
output:
  html_document:
    fig_height: 4.5
    highlight: pygments
    keep_md: yes
    number_sections: yes
    theme: readable
    toc: yes
---

# Introduction

A simple test of creating time based features to enhance classification performance. 

Basically, walking and running should be easy to differenciate by looking at acceleration variation on several sequencial samples.

Don't hesitate to comment / improve this quick try !


# Boilerplate 

Loading package and data.
Mainly :

* dplyr / tidyr for data manipulation
* ggplot2 / ggthemes / gridEXtra for plots
* MASS / CARET for Classificaton-Regression FW
  

```{r, results='hide', message=FALSE, warning=FALSE}
library('MASS')
library('caret')

library('ggplot2') # visualization
library('ggthemes') # visualization
library('scales') # visualization
library('plyr'); library('dplyr')
library('tidyr') # data manipulation
library('gridExtra')

library("ROCR")
library('stringr')
library('data.table')
library('zoo')

#library('DescTools')

setwd("~/repo/kaggle/walkOrRun/src/")

set.seed(2017)
select <- dplyr::select
options(digits.secs=6)
```

# Data overview

Load dataset :
```{r}
df <- fread('../input/dataset.csv', stringsAsFactors=TRUE)
df <- mutate(df, class = ifelse(df$activity == 0, "Walk", "Run")) %>%
      select(-activity, -date, -time, -username) %>% glimpse()
df$class <- as.factor(df$class)
```

Quick overview of some key parameters :
```{r}
df %>% 
  ggplot(aes(x=class, fill=class)) +
  geom_bar()
```

Small descriptive graph of parameters distribution
```{r}
df %>%
    gather( sensor, value, acceleration_x:gyro_z) %>%
    ggplot( aes(x = factor(class), y = value)) + 
    geom_jitter(alpha = 0.1) + geom_boxplot(color = "blue", outlier.colour = NA, fill = NA) +
    facet_wrap(~sensor)
```

## Simple feature separation caracterization : LDA

Preprocessing step :
  - Only scale here (TBC)

```{r}
intrain   <- createDataPartition(y=df$class , p=0.7, list=FALSE)
df_train  <- df[intrain,] ; df_test <- df[-intrain,]

df_train_PP <- preProcess(df_train[names(df_train) != "class"], method = c("scale"))
df_train_PP <- data.frame(predict(df_train_PP, df_train[names(df_train) != "class"]), class = df_train$class)
```


Simple feature $acceleration_x$ and $acceleration_y$ distribution graph per class :

```{r}
p1 <- ggplot(df_train_PP, aes(x = acceleration_x, y = acceleration_y, colour = factor(class))) + 
      geom_point()

p2 <- gather(df_train_PP, sensor, value, acceleration_x:acceleration_y) %>%
      ggplot(aes(x =value)) + geom_histogram(bins = 100) + 
      facet_grid(. ~ sensor, scales = "free_x")

grid.arrange(p1, p2)
```


LDA processing and differenciability :

```{r}
r       <- lda(formula = class ~ ., data = df_train_PP)
plda    <- predict(object = r,  newdata = df_train_PP)
dataset <- data.frame(class = df_train_PP[,"class"], lda = plda$x)
ggplot(dataset, aes(x=LD1,colour = class, fill = class)) + geom_density(alpha = 0.1)
```


# New Feature creation : Timeseries processing

Main idea is to process Data and Time to construct a GMT object, ordering sample by this GMT and compute a median absolute deviation (MAD) to caracterize variability of acceleration / gyros parameters.

```{r}
windowsSize = 5 
maxDeltaSecInWindow = 5

dftps <- fread('../input/dataset.csv', stringsAsFactors=FALSE)
dftps <- mutate(dftps, class = ifelse(dftps$activity == 0, "Walk", "Run"),
                secAsFloat =  str_replace(gsub("[0-9]+:[0-9]+:", "", time), ":", "."),
                gmt =  paste(date, gsub(":[0-9]+:[0-9]+$", "", time), sep='-'),
                gmt =  paste(gmt, secAsFloat, sep = ':') ,
                gmt =  as.POSIXct(strptime(gmt, "%Y-%m-%d-%H:%M:%OS"))) %>%
        select(-activity, -username, -secAsFloat, -date, -time) %>% arrange(gmt)

glimpse(dftps)
```

Determine valid time windows (ie. first gmt - last gmt < maxDeltaSecInWindow).
Here, we take 5 samples windows and min max GMT on this window must no exceed 5s (conservative).

```{r}
windowsSize = 5 
maxDeltaSecInWindow = 5
```

Perform rolling variance / rolling MAD on windowSize Timeframe
```{r}
dftps$mad_accel_x   <-  rollapply(dftps$acceleration_x, width=windowsSize, FUN=mad, align="center", partial=0)
dftps$mad_accel_y   <-  rollapply(dftps$acceleration_y, width=windowsSize, FUN=mad, align="center", partial=0)
dftps$mad_accel_z   <-  rollapply(dftps$acceleration_z, width=windowsSize, FUN=mad, align="center", partial=0)
dftps$mad_gyro_x    <-  rollapply(dftps$gyro_x, width=windowsSize, FUN=mad, align="center", partial=0)
dftps$mad_gyro_y    <-  rollapply(dftps$gyro_y, width=windowsSize, FUN=mad, align="center", partial=0)
dftps$mad_gyro_z    <-  rollapply(dftps$gyro_z, width=windowsSize, FUN=mad, align="center", partial=0)

dftps               <-  dftps %>%
                        mutate(validity = ifelse(abs(gmt - lag(gmt,windowsSize)) < maxDeltaSecInWindow, 1, 0),
                               mad_accel_x = ifelse( validity == 1, mad_accel_x, NA),
                               mad_accel_y = ifelse( validity == 1, mad_accel_y, NA),
                               mad_accel_z = ifelse( validity == 1, mad_accel_z, NA),
                               mad_gyro_x = ifelse( validity == 1, mad_gyro_x, NA),
                               mad_gyro_y = ifelse( validity == 1, mad_gyro_y, NA),
                               mad_gyro_z = ifelse( validity == 1, mad_gyro_z, NA))
glimpse(dftps)
```


## Simple feature separation caracterization for this timeseries derived feature

```{r}
intrain       <- createDataPartition(y=dftps$class , p=0.7, list=FALSE)
dftps_train   <- dftps[intrain,]  %>% select(-gmt, -validity)
dftps_test    <- dftps[-intrain,] %>% select(-gmt, -validity)

dftps_train_PP   <- preProcess(dftps_train[names(dftps_train) != "class"], method = c("scale"))
dftps_train_PP   <- data.frame(predict(dftps_train_PP, dftps_train[names(dftps_train) != "class"]), class = dftps_train$class)
```

Simple feature $acceleration_x$ and $acceleration_y$ distribution graph per class :

```{r}
p1 <- ggplot(dftps_train_PP, aes(x = mad_accel_x, y = mad_accel_y, colour = factor(class))) + 
      geom_point()

p2 <- gather(dftps_train_PP, sensor, value, mad_accel_x:mad_accel_y) %>%
      ggplot(aes(x =value)) + geom_histogram(bins = 100) + 
      facet_grid(. ~ sensor, scales = "free_x")

grid.arrange(p1, p2)
```


LDA processing and Run / Walk differenciability :

```{r waring='hide', message='hide'}
r       <- lda(formula = class ~ ., data = dftps_train_PP)
plda    <- suppressWarnings(predict(object = r,  newdata = dftps_train_PP))
```

Distribution plot :

```{r}
dataset <- data.frame(class = dftps_train_PP[,"class"], lda = plda$x)
ggplot(dataset, aes(x=LD1, colour = class, fill = class)) + geom_density(alpha = 0.1)
```

Look nice ;)

# Simple Model Fitting using temporal derived feature

Simple Logistic Regression  (extracted from https://www.kaggle.com/pcharambira/logistic-regression-vs-decision-tree on the same dataset)
```{r, eval=FALSE, include=FALSE}
trControl <- trainControl(method = "cv",  number=10, verboseIter = FALSE)
rf_fit <- caret::train(class ~   acceleration_x +
                               acceleration_y +
                               acceleration_z + 
                               mad_accel_x +        
                               mad_accel_y +
                               mad_accel_z +
                               mad_gyro_x + 
                               mad_gyro_y + 
                               mad_gyro_z + 
                               gyro_x +
                               gyro_y +
                               gyro_z , 
                data=dftps_train_PP, 
                method = "C5.0",
                preProcess = c("BoxCox", "center", "scale", "nzv"),
                tuneLength = 7, 
                trControl = trControl,
                metric = 'Accuracy',
                na.action = na.pass)
c_50fit
```

```{r}
model_logistic = glm(class ~   acceleration_x +
                               acceleration_y +
                               acceleration_z + 
                               mad_accel_x +        
                               mad_accel_y +
                               mad_accel_z +
                               mad_gyro_x + 
                               mad_gyro_y + 
                               mad_gyro_z + 
                               gyro_x +
                               gyro_y +
                               gyro_z ,
                               family=binomial(link='logit'), control = list(maxit = 50), data=dftps_train_PP)
print(summary(model_logistic))

# Apply the algorithm to the training sample
prediction_training = predict(model_logistic, dftps_train_PP, type = "response")
prediction_training = ifelse(prediction_training > 0.5, "Walk", "Run")
error = mean(prediction_training != dftps_train_PP$class)
print(paste('Model Accuracy after a split along the 0.5 probability',1-error))

# Get the ROC curve and the AUC
p = predict(model_logistic, dftps_train_PP, type="response")
pr = prediction(p, dftps_train_PP$class)
prf = performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf)

auc = performance(pr, measure = "auc")
auc = auc@y.values[[1]]
print(paste("General Model Accuracy", auc))
```
# To be done 

* Better filtering to find 'coherent' time window of variable duration (zoo application ?)
* Automatically determine window size (here, 5 sample length is arbitrary)
* Work on more complex time based feature (MAD is quite a simple one)
* Work harder on model fitting (here, simple copy / paste from https://www.kaggle.com/pcharambira/logistic-regression-vs-decision-tree)
