---
title: "Walk Or Run"
author: "NM"
date: "19 juillet 2017"
output:
  md_document:
    toc: yes
  html_document:
    fig_height: 4.5
    highlight: pygments
    keep_md: yes
    number_sections: yes
    theme: readable
    toc: yes
---

# Introduction

A simple test of creating time based features to enhance classification performance. 

Basically, walking and running should be easy to differenciate by looking at acceleration variation on several sequencial samples.

With a radial SVM kernel, classifier accuracy is given at 0.9961137

To be done :
* Better filtering to find 'coherent' time window of variable duration (zoo application ?)
* Automatically determine window size (here, 5 sample length is arbitrary)
* Work on more complex time based feature (MAD is quite a simple one)



# Boilerplate 

Loading package and data.
Mainly :

* dplyr / tidyr for data manipulation
* ggplot2 / ggthemes / gridEXtra for plots
* MASS / CARET for Classificaton-Regression FW
  

```{r, results='hide', message=FALSE, warning=FALSE}
library('MASS')
library('caret')

library('ggplot2') # visualization
library('ggthemes') # visualization
library('scales') # visualization
library('plyr'); library('dplyr')
library('tidyr') # data manipulation
library('gridExtra')

library('stringr')
library('data.table')
library('zoo')

setwd("~/repo/kaggle/walkOrRun/src/")

set.seed(2017)
select <- dplyr::select
options(digits.secs=6)
```

# Data overview

Load dataset :
```{r}
df <- fread('../input/dataset.csv', stringsAsFactors=TRUE)
df <- mutate(df, class = ifelse(df$activity == 0, "Walk", "Run")) %>%
      select(-activity, -date, -time, -username) %>% glimpse()
df$class <- as.factor(df$class)
```

Quick overview of some key parameters :
```{r}
df %>% 
  ggplot(aes(x=class, fill=class)) +
  geom_bar()
```

Small descriptive graph of parameters distribution
```{r}
df %>%
    gather( sensor, value, acceleration_x:gyro_z) %>%
    ggplot( aes(x = factor(class), y = value)) + 
    geom_jitter(alpha = 0.1) + geom_boxplot(color = "blue", outlier.colour = NA, fill = NA) +
    facet_wrap(~sensor)
```

## Simple feature separation caracterization : LDA

Preprocessing step :
  - Only scale here (TBC)

```{r}
intrain   <- createDataPartition(y=df$class , p=0.7, list=FALSE)
df_train  <- df[intrain,] ; df_test <- df[-intrain,]

df_train_PP <- preProcess(df_train[names(df_train) != "class"], method = c("scale"))
df_train_PP <- data.frame(predict(df_train_PP, df_train[names(df_train) != "class"]), class = df_train$class)
```


Simple feature $acceleration_x$ and $acceleration_y$ distribution graph per class :

```{r}
p1 <- ggplot(df_train_PP, aes(x = acceleration_x, y = acceleration_y, colour = factor(class))) + 
      geom_point()

p2 <- gather(df_train_PP, sensor, value, acceleration_x:acceleration_y) %>%
      ggplot(aes(x =value)) + geom_histogram(bins = 100) + 
      facet_grid(. ~ sensor, scales = "free_x")

grid.arrange(p1, p2)
```


LDA processing and differenciability :

```{r}
r       <- lda(formula = class ~ ., data = df_train_PP)
plda    <- predict(object = r,  newdata = df_train_PP)
dataset <- data.frame(class = df_train_PP[,"class"], lda = plda$x)
ggplot(dataset, aes(x=LD1,colour = class, fill = class)) + geom_density(alpha = 0.1)
```


# New Feature creation : Timeseries processing

Main idea is to process Data and Time to construct a GMT object, ordering sample by this GMT and compute a median absolute deviation (MAD) to caracterize variability of acceleration / gyros parameters.

```{r}
windowsSize = 5 
maxDeltaSecInWindow = 5

dftps <- fread('../input/dataset.csv', stringsAsFactors=FALSE)
dftps <- mutate(dftps, class = ifelse(dftps$activity == 0, "Walk", "Run"),
                secAsFloat =  str_replace(gsub("[0-9]+:[0-9]+:", "", time), ":", "."),
                gmt =  paste(date, gsub(":[0-9]+:[0-9]+$", "", time), sep='-'),
                gmt =  paste(gmt, secAsFloat, sep = ':') ,
                gmt =  as.POSIXct(strptime(gmt, "%Y-%m-%d-%H:%M:%OS"))) %>%
        select(-activity, -username, -secAsFloat, -date, -time) %>% arrange(gmt)

glimpse(dftps)
```

Determine valid time windows (ie. first gmt - last gmt < maxDeltaSecInWindow).
Here, we take 5 samples windows and min max GMT on this window must no exceed 5s (conservative).

```{r}
windowsSize = 5 
maxDeltaSecInWindow = 5
```

Perform rolling variance / rolling MAD on windowSize Timeframe
```{r}
dftps$mad_accel_x   <-  rollapply(dftps$acceleration_x, width=windowsSize, FUN=mad, align="center", partial=0)
dftps$mad_accel_y   <-  rollapply(dftps$acceleration_y, width=windowsSize, FUN=mad, align="center", partial=0)
dftps$mad_accel_z   <-  rollapply(dftps$acceleration_z, width=windowsSize, FUN=mad, align="center", partial=0)
dftps$mad_gyro_x    <-  rollapply(dftps$gyro_x, width=windowsSize, FUN=mad, align="center", partial=0)
dftps$mad_gyro_y    <-  rollapply(dftps$gyro_y, width=windowsSize, FUN=mad, align="center", partial=0)
dftps$mad_gyro_z    <-  rollapply(dftps$gyro_z, width=windowsSize, FUN=mad, align="center", partial=0)

dftps               <-  dftps %>%
                        mutate(temporalValidity = ifelse(abs(gmt - lag(gmt,windowsSize)) < maxDeltaSecInWindow, 1, 0),
                               temporalValidity = ifelse(is.na(temporalValidity), 0, temporalValidity))
glimpse(dftps)
```


## Simple feature separation caracterization for this timeseries derived feature

```{r}
intrain       <- createDataPartition(y=dftps$class , p=0.7, list=FALSE)
dftps_train   <- dftps[intrain,]  %>% select(-gmt)
dftps_test    <- dftps[-intrain,] %>% select(-gmt)

dftps_train_PP   <- preProcess(dftps_train[names(dftps_train) != "class"], method = c("scale"))
dftps_train_PP   <- data.frame(predict(dftps_train_PP, dftps_train[names(dftps_train) != "class"]), class = dftps_train$class)
```

Simple feature $acceleration_x$ and $acceleration_y$ distribution graph per class :

```{r}
p1 <- ggplot(dftps_train_PP, aes(x = mad_accel_x, y = mad_accel_y, colour = factor(class))) + 
      geom_point()

p2 <- gather(dftps_train_PP, sensor, value, mad_accel_x:mad_accel_y) %>%
      ggplot(aes(x =value)) + geom_histogram(bins = 100) + 
      facet_grid(. ~ sensor, scales = "free_x")

grid.arrange(p1, p2)
```


LDA processing and Run / Walk differenciability :

```{r waring='hide', message='hide'}
r       <- lda(formula = class ~ ., data = dftps_train_PP)
plda    <- suppressWarnings(predict(object = r,  newdata = dftps_train_PP))
```

Distribution plot :

```{r}
dataset <- data.frame(class = dftps_train_PP[,"class"], lda = plda$x)
ggplot(dataset, aes(x=LD1, colour = class, fill = class)) + geom_density(alpha = 0.1)
```

Look nice ;)

# Model Fitting using temporal derived feature

Simple SVM Radial training :
```{r}
trControl <- trainControl(method = "cv",  number=3, verboseIter = FALSE)
SVMRadial_fit <- caret::train(class ~   acceleration_x +
                               acceleration_y +
                               acceleration_z + 
                               mad_accel_x +        
                               mad_accel_y +
                               mad_accel_z +
                               mad_gyro_x + 
                               mad_gyro_y + 
                               mad_gyro_z + 
                               gyro_x +
                               gyro_y +
                               gyro_z , 
                data=dftps_train_PP, 
                method = "svmRadial",
                preProcess = c("center", "scale"),
                tuneLength = 5, 
                trControl = trControl,
                metric = 'Accuracy')
SVMRadial_fit
```
```{r}
# Performances:
p_SVMRad = predict(SVMRadial_fit, dftps_train_PP, type="raw")
print(confusionMatrix(p_SVMRad, dftps_train_PP$class))
postResample(p_SVMRad, dftps_train_PP$class)
```


# Classical GLM :
```{r}
model_logistic = caret::train(class ~   acceleration_x +
                               acceleration_y +
                               acceleration_z + 
                               mad_accel_x +        
                               mad_accel_y +
                               mad_accel_z +
                               mad_gyro_x + 
                               mad_gyro_y + 
                               mad_gyro_z + 
                               gyro_x +
                               gyro_y +
                               gyro_z ,
                               method="glmnet",
                               trControl = trControl,
                               data=dftps_train_PP)
print(summary(model_logistic))
```

```{r}
# Performances :
p_GLM = predict(model_logistic, dftps_train_PP, type = "raw")
print(confusionMatrix(p_GLM, dftps_train_PP$class))
postResample(p_GLM, dftps_train_PP$class)
```


